Using ReacTran

### 2015-10-25 notes/diffusion-local-times.Rmd
fixed bug in tran.1D fix

First, two quick checks: First, by setting $f_H=1$,
we should have the solution $u(x,t,r) = \exp(-tr)$.
Then, by setting $\sigma$ to be very small, the solution at $x$ should be $\exp(-trf_H(x))$.
```{r solve_pde_check, cache=TRUE, depends="solve_pde_setup"}
layout(t(1:2))
test.hetfn <- setup.prop.1D(grid=thegrid, func=function(x){rep(1,length(x))})
pde.fn <- function (t,y,parms,...) {
    tran <- tran.1D(C=y, A=clinefn, D=1/2, dx=thegrid)$dC
    list( tran - r * test.hetfn$mid * y )
}
pde.soln <- ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 )
matplot(pde.soln,type='l',ylab="probability of no recomb", col='black',log='y')
lines(tt,exp(-r*tt),lwd=2,lty=2,col='red')
## no diffusion
pde.fn <- function (t,y,parms,...) {
    tran <- tran.1D(C=y, A=clinefn, D=1e-6, dx=thegrid)$dC
    list( tran - r * hetfn$mid * y )
}
pde.soln <- ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 )
plot.x <- sort(match(sample(thegrid$x.mid[thegrid$x.mid>-5 & thegrid$x.mid<5],12),thegrid$x.mid))
matplot(pde.soln[,plot.x],type='l',ylab="probability of no recomb", col=rainbow(length(plot.x)), lty=1 )
matpoints( exp(outer( tt, -r*hetfn$mid[plot.x], "*" )), lty=4, col=rainbow(length(plot.x)), pch=20 )
```
**Looks good.**

Now, to solve the real problem:
```{r solve_pde, cache=TRUE, depends="solve_pde_setup"}
pde.fn <- function (t,y,parms,...) {
    tran <- tran.1D(C=y, A=clinefn, D=1/2, dx=thegrid)$dC
    list( tran - r * hetfn$mid * y )
}
pde.soln <- ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 ) # note FIRST COLUMN IS TIME
image(pde.soln,legend=TRUE,main="prob of no recomb by t", yaxt='n', ylab='space')
axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
```

@@

Actually, the cline gets set up
in a time-inhomogeneous way.
Let's do that, by numerical solution,
then run the lineage back through it.
The forwards-time evolution of the cline,
with isotropic density and migration,
solves the PDE (Barton 1979):
$$\begin{aligned}
  \partial_t p(x,t) = \frac{\sigma^2}{2} \partial_x^2 p(x,t)+ s p(x,t)(1-p(x,t))(2p(x,t)-1) .
\end{aligned}$$
```{r fwds_pde, cache=TRUE}
thegrid <- setup.grid.1D(x.up=-15, x.down=15, N = 5000)
yinit <- ifelse( thegrid$x.mid>0, 0, 1 )
tt <- seq(0,tau,length.out=100)
fwds.pde <- function (t,y,parms,...) {
    tran <- tran.1D(C=y, D=1/2, dx=thegrid)$dC
    list( tran + s * y * (1-y) * (2*y-1) )
}
fwds.soln <- ode.1D( y=yinit, times=tt, func=fwds.pde, nspec=1 ) # note FIRST COLUMN IS TIME
image(fwds.soln,legend=TRUE,main="mean cline development", yaxt='n', ylab='space')
axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
```
Now, let's run the killed lineage backwards in it:
```{r fwds_back_pde, cache=TRUE, depends="fwds_pde"}
yinit <- rep(1.0,thegrid$N)
rev.pde <- function (t,y,parms,...) {
    p <- cline_interp(tau-t,soln=fwds.soln)
    # p <- setup.prop.1D(grid=thegrid, xy=cbind(thegrid$x.mid,cline_interp(tau-t)), interp="linear")
    tran <- tran.1D(C=y, A=p, D=1/2, dx=thegrid)$dC
    list( tran - r * (1-p) * y )
}
rev.soln <- ode.1D( y=yinit, times=tt, func=rev.pde, nspec=1, tcrit=tau ) # note FIRST COLUMN IS TIME
image(rev.soln,legend=TRUE,main="prob of no recomb by t", yaxt='n', ylab='space')
axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
```

@@

Let's check this agrees with the solution for the PDE.
According to Knight 1981 (Thm 7.4.3), $\zeta'(0) = - 2 r \zeta(0)$;
ReacTran lets you specify the 'flux' at the boundary is proportional to $\zeta(0)$,
where the flux is in our case $\sigma^2/2 \zeta'$.
First note that the solution depends on where we put the boundaries:
```{r reflected_pde_boundaries,cache=TRUE}
# NOTE: must install ReacTran from R-forge, e.g. 
#   install.packages("ReacTran",repos="http://r-forge.r-project.org")
# or else
#   source("tran.1D.R")
# to fix bug in using convective boundary conditions.
x.eps <- 5*mean(thegrid$dx)
solve.reflected <- function (x.up=-20) {
    thegrid <- setup.grid.1D(x.up=x.up, x.down=0, N = 5000)
    hetfn <- setup.prop.1D(grid=thegrid, func=function(x){abs(x)<x.eps})
    yinit <- rep(1.0,thegrid$N)
    tau <- 40
    tt <- seq(0,tau,length.out=20)
    r <- 1
    pde.fn <- function (t,y,parms,...) {
        # flux = zeta'(0)/2 = a.bl.down*( zeta(0) - C.down )
        tran <- tran.1D(C=y, A=1, D=1/2, dx=thegrid, C.down=0, a.bl.down=4*r)$dC
        list( tran - r * hetfn$mid * y )
    }
    soln <- ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 ) # note FIRST COLUMN IS TIME
    attr(soln,"x") <- thegrid$x.mid
    return(soln)
}
pde.soln.0 <- solve.reflected(x.up=-5)
pde.soln.1 <- solve.reflected(x.up=-8)
pde.soln.2 <- solve.reflected(x.up=-20)
clineplot(pde.soln.0,x=attr(pde.soln.0,"x"),xlim=c(-5,0))
clineplot(pde.soln.1,x=attr(pde.soln.1,"x"),add=TRUE)
clineplot(pde.soln.2,x=attr(pde.soln.2,"x"),add=TRUE)
```

@@

Here's the cline shape:
```{r show_cline, cache=TRUE}
zeta.pcline <- function (x,r) { pcline(x,r,recombfn=function(x,r){zeta.reflected(x=x/sqrt(tau*sigma^2),r=r*tau)},tau=1,sigma=1) }
xx <- seq(-15,15,length.out=500)
for (tau in c(5,50)) {
    plot(xx, zeta.pcline(xx,r=1), type='n', 
            xlab="distance from cline center", ylab="prob inherited from the left", 
            main=as.expression(substitute(tau==mytau,list(mytau=tau))) )
    rr <- seq(0,0.2,length.out=11)
    pp <- sapply( rr, function (r) { zeta.pcline(xx,r=r,tau=tau) } )
    matplot(xx,pp,type='l',lty=1,col=rainbow(length(rr)),add=TRUE)
    legend("topright",lty=1,col=rainbow(length(rr)),legend=paste("r=",rr))
}
```

@@

Above we simulated the clines deriving from this process.
```{r sim_pushed, cache=TRUE}
rvals <- exp(seq(log(.001),log(0.5),length.out=10))
pushfn <- function(x,s){ifelse(x>0,-4*sqrt(s),0)}
nwalks <- 1e4
ntimes <- 1e3
s <- 0.1
tau <- 10
xinit <- seq(-8,8,length.out=nwalks)
sim.rw <- sim_diffusion( pushfn, tau, nwalks, ntimes, s=s, xinit=xinit )
sim.wt <- weighted_time( sim.rw, function (x) { x>0 } )
# distribution of time above zero depending on starting point
plot(xinit,sim.wt,xlab="initial position", ylab="time above zero")
```
We can also get the predicted clines by solving the PDE.
Note that here we are approximating $p(x) = \min(1,\exp(-4x\sqrt{s}))$.
```{r pushed_pde,cached=TRUE}
thegrid <- setup.grid.1D(x.up=-15, x.down=15, N = 5000)
clinefn <- setup.prop.1D(grid=thegrid, func=function(x,s){ifelse(x>0,exp(-4*x*sqrt(s)),1.0)}, s=s)
hetfn <- setup.prop.1D(grid=thegrid, func=function(x){ifelse(x>0,1,0)})
yinit <- rep(1.0,thegrid$N)
tt <- seq(0,tau,length.out=100)
pde.solns <- lapply( rvals, function (r) {
        pde.fn <- function (t,y,parms,...) {
            tran <- tran.1D(C=y, A=clinefn, D=1/2, dx=thegrid)$dC
            if (any(is.na(tran))) { browser() }
            list( tran - r * hetfn$mid * y )
        }
        ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 ) # note FIRST COLUMN IS TIME
      } )
layout(t(1:2))
for (k in c(6,10)) {
    clineplot(pde.solns[[k]],main=sprintf("r=%0.2f",tau,rvals[k]))
}
```

@@

And, finally, here's some actual clines predicted by this model:
```{r pushed_cline,cache=TRUE,depends="pushed_pde"}
tau <- 100
tt <- seq(0,tau,length.out=100)
pde.solns <- lapply( rvals, function (r) {
        pde.fn <- function (t,y,parms,...) {
            tran <- tran.1D(C=y, A=clinefn, D=1/2, dx=thegrid)$dC
            if (any(is.na(tran))) { browser() }
            list( tran - r * hetfn$mid * y )
        }
        ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 ) # note FIRST COLUMN IS TIME
      } )
plot.tvals <- floor(seq(0,tau,length.out=7))[-1]
layout(t(1:2))
for (tval in plot.tvals) {
    matplot( thegrid$x.mid, 
            pcline(thegrid$x.mid, p=sapply( pde.solns, cline_interp, T=tval ) ), 
                   type='l', main=sprintf("t=%d",tval) 
        )
    legend("bottomleft",lty=1:5,col=1:6,legend=paste("r=",round(rvals/2,digits=2)))
}
```

>>>>

First, two quick checks: First, by setting $f_H=1$,
we should have the solution $u(x,t,r) = \exp(-tr)$.
Then, by setting $\sigma$ to be very small, the solution at $x$ should be $\exp(-trf_H(x))$.
```{r solve_pde_check, cache=TRUE, depends="solve_pde_setup"}
layout(t(1:2))
test.hetfn <- setup.prop.1D(grid=thegrid, func=function(x){rep(1,length(x))})
pde.fn <- function (t,y,parms,...) {
    tran <- tran.1D(C=y, A=clinefn, D=1/2, flux.up=0, flux.down=0, dx=thegrid)$dC
    list( tran - r * test.hetfn$mid * y )
}
pde.soln <- ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 )
matplot(pde.soln,type='l',ylab="probability of no recomb", col='black',log='y')
lines(tt,exp(-r*tt),lwd=2,lty=2,col='red')
## no diffusion
pde.fn <- function (t,y,parms,...) {
    tran <- tran.1D(C=y, A=clinefn, D=1e-6, flux.up=0, flux.down=0, dx=thegrid)$dC
    list( tran - r * hetfn$mid * y )
}
pde.soln <- ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 )
plot.x <- sort(match(sample(thegrid$x.mid[thegrid$x.mid>-5 & thegrid$x.mid<5],12),thegrid$x.mid))
matplot(pde.soln[,plot.x],type='l',ylab="probability of no recomb", col=rainbow(length(plot.x)), lty=1 )
matpoints( exp(outer( tt, -r*hetfn$mid[plot.x], "*" )), lty=4, col=rainbow(length(plot.x)), pch=20 )
```
**Looks good.**

Now, to solve the real problem:
```{r solve_pde, cache=TRUE, depends="solve_pde_setup"}
pde.fn <- function (t,y,parms,...) {
    tran <- tran.1D(C=y, A=clinefn, D=1/2, flux.up=0, flux.down=0, dx=thegrid)$dC
    list( tran - r * hetfn$mid * y )
}
pde.soln <- ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 ) # note FIRST COLUMN IS TIME
image(pde.soln,legend=TRUE,main="prob of no recomb by t", yaxt='n', ylab='space')
axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
```

@@

Actually, the cline gets set up
in a time-inhomogeneous way.
Let's do that, by numerical solution,
then run the lineage back through it.
The forwards-time evolution of the cline,
with isotropic density and migration,
solves the PDE (Barton 1979):
$$\begin{aligned}
  \partial_t p(x,t) = \frac{\sigma^2}{2} \partial_x^2 p(x,t)+ s p(x,t)(1-p(x,t))(2p(x,t)-1) .
\end{aligned}$$
```{r fwds_pde, cache=TRUE}
thegrid <- setup.grid.1D(x.up=-15, x.down=15, N = 5000)
yinit <- ifelse( thegrid$x.mid>0, 0, 1 )
tt <- seq(0,tau,length.out=100)
fwds.pde <- function (t,y,parms,...) {
    tran <- tran.1D(C=y, D=1/2, dx=thegrid)$dC
    list( tran + s * y * (1-y) * (2*y-1) )
}
fwds.soln <- ode.1D( y=yinit, times=tt, func=fwds.pde, nspec=1 ) # note FIRST COLUMN IS TIME
image(fwds.soln,legend=TRUE,main="mean cline development", yaxt='n', ylab='space')
axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
```
Now, let's run the killed lineage backwards in it:
```{r fwds_back_pde, cache=TRUE, depends="fwds_pde"}
yinit <- rep(1.0,thegrid$N)
rev.pde <- function (t,y,parms,...) {
    p <- cline_interp(tau-t,soln=fwds.soln)
    # p <- setup.prop.1D(grid=thegrid, xy=cbind(thegrid$x.mid,cline_interp(tau-t)), interp="linear")
    tran <- tran.1D(C=y, A=p, D=1/2, flux.up=0, flux.down=0, dx=thegrid)$dC
    list( tran - r * (1-p) * y )
}
rev.soln <- ode.1D( y=yinit, times=tt, func=rev.pde, nspec=1, tcrit=tau ) # note FIRST COLUMN IS TIME
image(rev.soln,legend=TRUE,main="prob of no recomb by t", yaxt='n', ylab='space')
axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
```

@@

Let's check this agrees with the solution for the PDE.
According to Knight 1981 (Thm 7.4.3), $\zeta'(0) = - 2 r \zeta(0)$;
ReacTran lets you specify the 'flux' at the boundary is proportional to $\zeta(0)$,
where the flux is in our case $\sigma^2/2 \zeta'$.
First note that the solution depends on where we put the boundaries:
```{r reflected_pde_boundaries,cache=TRUE}
# NOTE: must install ReacTran from R-forge, e.g. 
#   install.packages("ReacTran",repos="http://r-forge.r-project.org")
# or else
source("tran.1D.R")
# to fix bug in using convective boundary conditions.
x.eps <- 5*mean(thegrid$dx)
solve.reflected <- function (x.up=-20) {
    thegrid <- setup.grid.1D(x.up=x.up, x.down=0, N = 5000)
    hetfn <- setup.prop.1D(grid=thegrid, func=function(x){abs(x)<x.eps})
    yinit <- rep(1.0,thegrid$N)
    tau <- 40
    tt <- seq(0,tau,length.out=20)
    r <- 1
    pde.fn <- function (t,y,parms,...) {
        # flux = zeta'(0)/2 = a.bl.down*( zeta(0) - C.down )
        tran <- tran.1D(C=y, A=1, D=1/2, dx=thegrid, C.down=0, a.bl.down=4*r)$dC
        list( tran - r * hetfn$mid * y )
    }
    soln <- ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 ) # note FIRST COLUMN IS TIME
    attr(soln,"x") <- thegrid$x.mid
    return(soln)
}
pde.soln.0 <- solve.reflected(x.up=-5)
pde.soln.1 <- solve.reflected(x.up=-8)
pde.soln.2 <- solve.reflected(x.up=-20)
clineplot(pde.soln.0,x=attr(pde.soln.0,"x"),xlim=c(-5,0))
clineplot(pde.soln.1,x=attr(pde.soln.1,"x"),add=TRUE)
clineplot(pde.soln.2,x=attr(pde.soln.2,"x"),add=TRUE)
```

@@

Here's the cline shape:
```{r show_cline, cache=TRUE}
zeta.pcline <- function (x,r) { pcline(x,r,recombfn=function(x,r,...){zeta.reflected(x=x/sqrt(tau*sigma^2),r=r*tau)},tau=1,sigma=1) }
xx <- seq(-15,15,length.out=500)
for (tau in c(5,50)) {
    plot(xx, zeta.pcline(xx,r=1), type='n', 
            xlab="distance from cline center", ylab="prob inherited from the left", 
            main=as.expression(substitute(tau==mytau,list(mytau=tau))) )
    rr <- seq(0,0.2,length.out=11)
    pp <- sapply( rr, function (r) { zeta.pcline(xx,r=r,tau=tau) } )
    matplot(xx,pp,type='l',lty=1,col=rainbow(length(rr)),add=TRUE)
    legend("topright",lty=1,col=rainbow(length(rr)),legend=paste("r=",rr))
}
```

@@

Above we simulated the clines deriving from this process.
```{r sim_pushed, cache=TRUE}
rvals <- exp(seq(log(.001),log(0.5),length.out=10))
pushfn <- function(x,s){ifelse(x>0,-4*sqrt(s),0)}
nwalks <- 1e4
ntimes <- 1e3
s <- 0.1
tau <- 10
xinit <- seq(-8,8,length.out=nwalks)
sim.rw <- sim_diffusion( pushfn, tau, nwalks, ntimes, s=s, xinit=xinit )
sim.wt <- weighted_time( sim.rw, function (x) { x>0 } )
# distribution of time above zero depending on starting point
plot(xinit,sim.wt,xlab="initial position", ylab="time above zero")
```
We can also get the predicted clines by solving the PDE.
Note that here we are approximating $p(x) = \min(1,\exp(-4x\sqrt{s}))$.
```{r pushed_pde,cached=TRUE}
thegrid <- setup.grid.1D(x.up=-15, x.down=15, N = 5000)
clinefn <- setup.prop.1D(grid=thegrid, func=function(x,s){ifelse(x>0,exp(-4*x*sqrt(s)),1.0)}, s=s)
hetfn <- setup.prop.1D(grid=thegrid, func=function(x){ifelse(x>0,1,0)})
yinit <- rep(1.0,thegrid$N)
tt <- seq(0,tau,length.out=100)
pde.solns <- lapply( rvals, function (r) {
        pde.fn <- function (t,y,parms,...) {
            tran <- tran.1D(C=y, A=clinefn, D=1/2, flux.up=0, flux.down=0, dx=thegrid)$dC
            if (any(is.na(tran))) { browser() }
            list( tran - r * hetfn$mid * y )
        }
        ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 ) # note FIRST COLUMN IS TIME
      } )
layout(t(1:2))
for (k in c(6,10)) {
    clineplot(pde.solns[[k]],main=sprintf("r=%0.2f",tau,rvals[k]))
}
```

@@

And, finally, here's some actual clines predicted by this model:
```{r pushed_cline,cache=TRUE,depends="pushed_pde"}
tau <- 100
tt <- seq(0,tau,length.out=100)
pde.solns <- lapply( rvals, function (r) {
        pde.fn <- function (t,y,parms,...) {
            tran <- tran.1D(C=y, A=clinefn, D=1/2, flux.up=0, flux.down=0, dx=thegrid)$dC
            if (any(is.na(tran))) { browser() }
            list( tran - r * hetfn$mid * y )
        }
        ode.1D( y=yinit, times=tt, func=pde.fn, nspec=1 ) # note FIRST COLUMN IS TIME
      } )
plot.tvals <- floor(seq(0,tau,length.out=7))[-1]
layout(t(1:2))
for (tval in plot.tvals) {
    matplot( thegrid$x.mid, 
            pcline(thegrid$x.mid, p=sapply( pde.solns, cline_interp, T=tval ) ), 
                   type='l', main=sprintf("t=%d",tval) 
        )
    legend("bottomleft",lty=1:5,col=1:6,legend=paste("r=",round(rvals/2,digits=2)))
}
```

### 2015-10-25 notes/pde-clines.Rmd
fixed bug in tran.1D fix

Suppose that a lineage moves in a frequency profile $p(x)$ of allele $A$,
switching between backgrounds at rate $r p(x)$ or $r (1-p(x))$.
Let $(X_t,Z_t)$ denote this process,
so that
$$\begin{aligned}
    dX_t &= \grad \log(p(X_t)) dt + dB_t  \qquad \text{if } Z_t = A \\
         &= \grad \log(1-p(X_t)) dt + dB_t  \qquad \text{if } Z_t = B \\
    \P\{ Z_{t+\epsilon} = B | Z_t = A \} &= \epsilon r (1-p(X_t)) + O(\epsilon^2) \\
    \P\{ Z_{t+\epsilon} = A | Z_t = B \} &= \epsilon r p(X_t) + O(\epsilon^2) .
\end{aligned}$$
(Note that $p$ may also depend on $t$.)
We would like to find the probability that 
a lineage which we sample at location $x$ on background $z$,
derived from background $A$ $\tau$ units of time ago,
which we denote by
$$
q(x,z,\tau,r) = \P\{ Z_\tau = A \} .
$$
(Note that this gives us linkage to the selected site,
and for the cline shape we only need $p(x,\tau) q(x,A,\tau,r) + (1-p(x,\tau) q(x,B,\tau,r)$,
but it is easier to compute this.)

@@

Let's solve this numerically.
```{r solve_pde, cache=TRUE, depends="solve_pde_setup"}
r <- 0.1
s <- 0.1
sigma <- 10
zone.age <- 100  # this is tau
tt <- seq(0,zone.age,length.out=100)
pde.soln <- solve.pde( u=log_u, um1=log_1mu, r=r, times=tt, 
                  s=s, sigma=sigma, grid=thegrid, log.u=TRUE )
image(pde.soln,legend=TRUE,main=sprintf("q for z=%s, r=%0.3f",c("A","B"),r), yaxt='n', ylab='space',zlim=c(0,1))
axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
```
These combine to give the predicted cline:
```{r show_pde_cline,cache=TRUE,depends=c("solve_pde_setup","solve_pde")}
pde.cline <- cline.from.soln(pde.soln,grid=thegrid)
image(pde.cline,legend=TRUE,zlim=c(0,1),main=sprintf("clines, at r=%0.2f",r))
# same thing, plotted with lines
clineplot(pde.cline,main=sprintf("clines, at r=%0.2f",r))
```
**Note:** the range should be wide enough these are *going to zero*
at the ends, or else we have to worry about the boundary conditions.

Comparison to simulation
------------------------

@@

Here's the PDE solution:
```{r match_pde,cache=TRUE,depends="solve_pde_setup"}
tt <- seq(0,zone.age,length.out=100)
pde.solns <- lapply(rvals, function (r) { solve.pde( u=log_u, um1=log_1mu, r=r, times=tt, 
                                                s=s, sigma=sigma, grid=thegrid, log.u=TRUE ) } )
pde.clines <- lapply( pde.solns, cline.from.soln, grid=thegrid )
match.clines <- sapply(pde.clines,function(x)x[length(tt),-1])
```
Now, comparing the PDE solution to the others:
```{r match_cline_pde}
for (k in seq_along(rvals)) {
    plot(xx,freqs[,k],type='l',xlab="spatial location",ylab="proportion",main=sprintf("r=%0.3f",rvals[k]))
    lines(xx,1-tanh_cline(xx,s=s,sigma=sigma),col='red',lwd=3,lty=3)
    lines(1-thegrid$x.mid,match.clines[,k],col='green',lwd=3)
}
```

>>>>

Suppose that a lineage moves in a frequency profile $p(x)$ of allele $A$,
switching between backgrounds at rate $r p(x)$ or $r (1-p(x))$.
Let $(X_t,Z_t)$ denote this process,
so that
$$\begin{aligned}
    dX_t &= \grad \log(p(X_t)) dt + dB_t  \qquad \text{if } Z_t = A \\
         &= \grad \log(1-p(X_t)) dt + dB_t  \qquad \text{if } Z_t = B \\
    \P\{ Z_{t+\epsilon} = B | Z_t = A \} &= \epsilon r (1-p(X_t)) + O(\epsilon^2) \\
    \P\{ Z_{t+\epsilon} = A | Z_t = B \} &= \epsilon r p(X_t) + O(\epsilon^2) .
\end{aligned}$$
(Note that $p$ may also depend on $t$.)
We would like to find the probability that 
a lineage which we sample at location $x$ on background $z$,
derived from background $A$ $\tau$ units of time ago,
which we denote by
$$
q(x,z,\tau,r) = \P^x\{ Z_\tau = A \} .
$$
(Note that this gives us linkage to the selected site,
and for the cline shape we only need $p(x,\tau) q(x,A,\tau,r) + (1-p(x,\tau))q(x,B,\tau,r)$,
but it is easier to compute this.)
Also note that by symmetry, we should have that
if $r>0$, then
$$\begin{aligned}
    q(x,z,\tau,r) \to \frac{1}{2} \qquad \text{as } \tau \to \infty .
\end{aligned}$$

@@

Let's solve this numerically.
```{r solve_pde, cache=TRUE, depends="solve_pde_setup"}
r <- 0.1
s <- 0.1
sigma <- 10
zone.age <- 200  # this is tau
tt <- seq(0,zone.age,length.out=100)
pde.soln <- solve.pde( u=log_u, um1=log_1mu, r=r, times=tt, 
                  s=s, sigma=sigma, grid=thegrid, log.u=TRUE )
image(pde.soln,legend=TRUE,main=sprintf("q for z=%s, r=%0.3f",c("A","B"),r), yaxt='n', ylab='space',zlim=c(0,1))
axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
```
These combine to give the predicted cline:
```{r show_pde_cline,cache=TRUE,depends=c("solve_pde_setup","solve_pde")}
pde.cline <- cline.from.soln(pde.soln,grid=thegrid)
image(pde.cline,legend=TRUE,zlim=c(0,1),main=sprintf("clines, at r=%0.2f",r))
# same thing, plotted with lines
clineplot(pde.cline,x=thegrid$x.mid,main=sprintf("clines, at r=%0.2f, up to t=%d",r,zone.age))
```
**Note:** the range should be wide enough these are *going to zero*
at the ends, or else we have to worry about the boundary conditions.

The smaller $r$ is, the slower the per-background probabilities smooth out:
```{r small_r_sim}
r <- 0.01
pde.soln <- solve.pde( u=log_u, um1=log_1mu, r=r, times=tt, 
                  s=s, sigma=sigma, grid=thegrid, log.u=TRUE )
image(pde.soln,legend=TRUE,main=sprintf("q for z=%s, r=%0.3f",c("A","B"),r), yaxt='n', ylab='space',zlim=c(0,1))
axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
pde.cline <- cline.from.soln(pde.soln,grid=thegrid)
clineplot(pde.cline,x=thegrid$x.mid,main=sprintf("clines, at r=%0.2f, up to t=%d",r,zone.age))
```

And, the cline flattens out *very* slowly:
```{r longer_t_sim}
r <- 0.1
zone.age <- 2000  # this is tau
tt <- seq(0,zone.age,length.out=100)
pde.soln <- solve.pde( u=log_u, um1=log_1mu, r=r, times=tt, 
                  s=s, sigma=sigma, grid=thegrid, log.u=TRUE )
image(pde.soln,legend=TRUE,main=sprintf("q for z=%s, r=%0.3f",c("A","B"),r), yaxt='n', ylab='space',zlim=c(0,1))
axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
pde.cline <- cline.from.soln(pde.soln,grid=thegrid)
clineplot(pde.cline,x=thegrid$x.mid,main=sprintf("clines, at r=%0.2f, up to t=%d",r,zone.age))
```
This should be symmetric.  Let's check:
```{r check_symmetry}
pde.soln <- solve.pde( um1=log_u, u=log_1mu, r=r, times=tt, 
                yinit=c( rep(1.0,thegrid$N), rep(0.0,thegrid$N) ),
                s=s, sigma=sigma, grid=thegrid, log.u=TRUE )
image(pde.soln,legend=TRUE,main=sprintf("q for z=%s, r=%0.3f",c("A","B"),r), yaxt='n', ylab='space',zlim=c(0,1))
axis(2,at=(pretty(thegrid$x.mid)-min(thegrid$x.mid))/diff(range(thegrid$x.mid)),labels=pretty(thegrid$x.mid))
pde.cline <- cline.from.soln(pde.soln,grid=thegrid)
clineplot(pde.cline,x=thegrid$x.mid,main=sprintf("clines, at r=%0.2f, up to t=%d",r,zone.age))
```

Comparison to simulation
========================

@@

Here's the PDE solution:
```{r match_pde,cache=TRUE,depends="solve_pde_setup"}
tt <- seq(0,zone.age,length.out=100)
pde.solns <- lapply(rvals, function (r) { solve.pde( u=log_u, um1=log_1mu, r=r, times=tt, 
                                                s=s, sigma=sigma, grid=thegrid, log.u=TRUE ) } )
pde.clines <- lapply( pde.solns, cline.from.soln, grid=thegrid )
match.clines <- sapply(pde.clines,function(x)x[length(tt),-1])
for (k in floor(seq(1,length(rvals),length.out=3))) {
    image(pde.solns[[k]],legend=TRUE,main=sprintf("r=%f",rvals[k]))
}
```
Now, comparing the PDE solution to the others:
```{r match_cline_pde}
for (k in seq_along(rvals)) {
    plot(xx,freqs[,k],type='l',xlab="spatial location",ylab="proportion",main=sprintf("r=%0.3f",rvals[k]))
    lines(xx,1-tanh_cline(xx,s=s,sigma=sigma),col='red',lwd=3,lty=3)
    lines(1-thegrid$x.mid,match.clines[,k],col='green',lwd=3)
}
```

### 2015-10-25 notes/reactran-notes.Rmd
fixed bug in tran.1D fix

@@

>>>>

The interface area
------------------

@@

<!--
Boundary conditions
-------------------

We want the function `tran.1D` to return the generator of a Markov process
applied to a function
-- not a sub-Markov process, with killing,
as that appears in the reaction term.
What should the boundary conditions be?
Let $G$ be the generator,
which we write in the form
$$\begin{aligned}
Gf = \frac{1}{A} \grad \cdot A D \grad f,
\end{aligned}$$
where $D \grad f$ is the *flux*.
We know that $G\textbf{1}=0$,
where $\textbf{1}$ is the constant function.
Well,
$$\begin{aligned}
    G\textbf{1} 
    &=
\end{aligned}$$

-->
